---
title: "Merge"
editor: source
---


```{r, echo = FALSE}
source("./R/init.R")
log_info("Called git-merge notebook")
```

## Recap

From the previous branching example - 

1. We delivered stage one of an analysis to the client
2. We started on the secondary analyses 
3. Before the secondary analysis was complete, we realised there was an error in the original analysis that needed an emergency fix
3. We rewound to an earlier state in the repository and then fixed the error and checked in our work
4. We made the work permanent by creating a new branch and then checking out that branch
5. We tagged the fix and the re-issued the analysis to the client
6. We jumped back to our secondary analysis by checking out the main branch

Ok, so now we have a fix to the error in the original analysis in one branch `fix-01` and a partially completed secondary analysis on the `main` branch.
We want to bring the changes from the fix into our present work.

## Merge processes

There are two related approaches for combining your changes into a shared repository; rebase and merge.
Personally, I never use `git rebase` and always use `git merge`.
Mainly this is because the complexity of the repositories that I work on never justifies the rebase approach.
Rebasing rewrites commits from one branch to another whereas merge retains the full repository history.

Merging is usually a fairly automated process.
Run the following.

```{bash, eval = FALSE, engine = "sh"}
git checkout main
git merge fix-01
```

You may be prompted to enter a commit message (alternatively just provide the -m flag).

```{bash, eval = FALSE, engine = "sh"}
Merge branch 'fix-01'
# Please enter a commit message to explain why this merge is necessary,
# especially if it merges an updated upstream into a topic branch.
#
# Lines starting with '#' will be ignored, and an empty message aborts
# the commit.
```

and then the following output.

```{bash, eval = FALSE, engine = "sh"}
## Auto-merging branching.R
## Merge made by the 'ort' strategy.
##  branching.R | 3 ++-
##  1 file changed, 2 insertions(+), 1 deletion(-)
```

If you look at the difference between the HEAD (the `HEAD~1` notation means compare with the previous commit, `HEAD~2` means compare with 2 commits prior and so on) and the commit associated with the merge you will see

```{bash, eval = FALSE, engine = "sh"}
git diff HEAD~1
## diff --git a/branching.R b/branching.R
## index b60c9f0..8f1c1b3 100644
## --- a/branching.R
## +++ b/branching.R
## @@ -45,6 +45,7 @@ message("\nDESCRIPTIVE SUMMARY:\n")
## 
##  lyt <- basic_table() %>%
##    split_cols_by("arm") %>%
## +  split_rows_by("age") %>%
##    summarize_row_groups() %>%
##    analyze("y", mean, format = "xx.x")
## 
## @@ -54,7 +55,7 @@ build_table(lyt, d)
## 
##  message("\n\nANALYSIS OF CONTINOUS OUTCOME (UNSTRATIFIED):\n")
## 
## -lm1 <- lm(y ~ x, data = d)
## +lm1 <- lm(y ~ x * u, data = d)
##  summary(lm1)
## 
##  message("\n\nANALYSIS OF BINARY OUTCOME (UNSTRATIFIED):\n")
```

We now have the changes from the emergency fix in the main branch and we can continue with the secondary analysis.

Sometimes merging doesn't work quite so smoothly and we need to iron out conflicts.

## Exercises

:::{#exr-1} 

Complete the analysis by adding the following code to generate a figure from the fitted model.

```{r, eval = FALSE}
p1 <- ggplot(d_fit, aes(x = arm, y  = pr_z)) +
  geom_point() +
  geom_linerange(aes(ymin = pr_z_lb, ymax = pr_z_ub)) +
  scale_x_discrete("") +
  scale_y_continuous("Probability of response")

ggsave("fig-sec.png", p1, width = 10, height = 10, units = "cm")
```

Here is what you need to do:

- Run the updated script `Rscript branching.R` to make sure it works.
- Edit the `.gitignore` file so that the figure does not get committed to the repository.
- Stage and commit the files and then review the commit history.
- Create a v2.0 tag with an meaningful message.
:::


:::{#exr-2} 

Usually it makes sense to create a new branch for each piece of development we undertake.
This ensures that the main branch continues to reflect a working version at all times.
Create a new branch from the current state and call it `analysis-03`.
Run the following code:

```{bash, eval = F, engine = "sh"}
git branch analysis-03
git checkout analysis-03
git status
```

Add the analysis code into `branching.R`

```{r, eval = FALSE}
message("\n\nANALYSIS OF SURVIVAL OUTCOME (UNSTRATIFIED):\n")
lm3 <- coxph(Surv(w, evt) ~ x, data = d)
summary(lm3)
```

Run the script to make sure it works then stage and commit along with an updated `.gitignore` files that excludes all `png` files.

Checkout main:

```{bash, eval = F, engine = "sh"}
git checkout main
```

Add the following change to the `branching.R` code (just adding a new theme to the ggplot figure)

```{r, eval = FALSE}
p1 <- ggplot(d_fit, aes(x = arm, y  = pr_z)) +
  geom_point() +
  geom_linerange(aes(ymin = pr_z_lb, ymax = pr_z_ub)) +
  scale_x_discrete("") +
  scale_y_continuous("Probability of response") +
  theme_bw()
```

Run the script, then stage and commit.

Checkout `analysis-03` branch:

```{bash, eval = F, engine = "sh"}
git checkout analysis-03
```

Add the following code:

```{r, eval = FALSE}
png("fig-surv.png")
plot(survfit(Surv(w, evt) ~ x, data = d), lty = 1:2)
dev.off()
```

Run the script to make sure it works then stage and commit.
Treat this as the release by tagging it as `v3.0`.
Checkout the `main` branch and merge the analysis into main.

```{bash, eval = F, engine = "sh"}
git checkout main
git merge analysis-03
```

Review the commit history:

```{bash, eval = F, engine = "sh"}
git log --oneline --decorate --graph --all
## *   2829471 (HEAD -> main) Merge branch 'analysis-03'
## |\
## | * be09457 (tag: v3.0, analysis-03) Finished surv
## | * 7a5e7b9 Surv analysis
## * | 53115b6 minor
## |/
## * 1a9dfeb (tag: v2.0) Edits from code review
## * d1586df Completion of secondary analysis
## *   bf34b9c Merge from fix-01
## |\
## | * b8de16c (tag: v1.1, fix-01) Edits from code review
## | * 12561d9 Emergency fix
## * | d88f9e5 WIP
## |/
## * a2cc6f7 (tag: v1.0) Comments from code review
## * fa24778 branching.R
## * 3bdac46 Revised approach to capturing input
## * 327170a Minor edit
## * b078716 New file
## * 0cd2d52 First commit
```
:::

At this point, you have the fundamentals.
However, there are things that are going to catch you out.
For example, in the above work, we used three-way merges.
These are convenient but can also get messy when your branching structure gets complex.
In most of our daily work merges will probably suffice, but know that there is another way.
The other way is rebasing, but we will not deal with it here.
